# Preface {.unnumbered}

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = NA, fig.align = "center", fig.width = 4, fig.height = 4, message = FALSE, warning = FALSE)
library(tidyverse)
# Parallel Processing
library(doMC)
registerDoMC(cores = 16)
```

This material is from the [DataCamp](https://www.datacamp.com) course [Modeling with Data in the Tidyverse](https://learn.datacamp.com/courses/modeling-with-data-in-the-tidyverse) by Albert Y. Kim.

**Course Description:** In this course, you will learn to model with data. Models attempt to capture the relationship between an outcome variable of interest and a series of explanatory/predictor variables. Such models can be used for both explanatory purposes, e.g. "Does knowing professors' ages help explain their teaching evaluation scores?", and predictive purposes, e.g., "How well can we predict a house's price based on its size and condition?" You will leverage your tidyverse skills to construct and interpret such models. This course centers around the use of linear regression, one of the most commonly-used and easy to understand approaches to modeling. Such modeling and thinking is used in a wide variety of fields, including statistics, causal inference, machine learning, and artificial intelligence.

Reminder to self: each `*.qmd` file contains one and only one chapter, and a chapter is defined by the first-level heading `#`.

```{r include=FALSE, eval = TRUE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown', 'caret', 
  'caTools', 'ggplot2', "ranger", "randomForest", 'tidyverse',
  'broom', 'dplyr', 'pwr', 'haven', 'simputation', 'cluster', 
  'agricolae', 'naniar', 'mice', 'moderndive'), 
  'packages.bib')
```
